---
title: "Inference"
author: "Jason Luo"
date: "2025-01-02"
output: pdf_document
---

```{r, message = FALSE, warning=FALSE}
rm(list = ls())

setwd('/Users/jasonluo/Documents/R_crime_analysis')
library(tidyverse)

# Importing cleaned data
df <- read.csv('RMS_Crime_Incidents_Cleaned.csv')

unique(df$year)
```

## ${\chi}^2$ test 
We run a ${\chi}^2$ test for Homogeneity, testing if there is a difference in the amount of crimes that occur in each zipcode per year
``` {r}

get_num_crimes_by_zip_year <- function(data, yr) {
  
  
  print(paste0("Year: ", yr))
  crimes <- data %>%
    select(year, zip_code) %>%
    filter(year == yr) %>%
    group_by(zip_code) %>%
    summarise(num_crimes = n())
  
  return(crimes)
}


num_crimes_by_zip_year <- list()
years = c(2021,2022,2023)

for (i in seq_along(years)) {
  crimes <- get_num_crimes_by_zip_year(df, years[i]) 
  num_crimes_by_zip_year[[i]] <- crimes$num_crimes
}

```

``` {r}
zip_codes <- crimes$zip_code # all 3 years have the same zip codes 
contingency_table <- rbind(num_crimes_by_zip_year[[1]], num_crimes_by_zip_year[[2]], num_crimes_by_zip_year[[3]])
dimnames(contingency_table) <- list(Year = years, Zipcode = zip_codes)

contingency_table
```
Displaying the ${\chi}^2$ test results
``` {r}
Xsq <- chisq.test(contingency_table)
Xsq
```

Other related quantities:
``` {r}
Xsq$observed   # observed counts
Xsq$expected   # expected counts under the null
Xsq$residuals  # Pearson residuals
Xsq$stdres     # standardized residual
```
## Poisson Regression
Using poisson regression to model number (count) of crime incidents based off of neighborhood,
council district, year of occurence, and zip code

``` {r}

df_agg <- df %>%
  group_by(zip_code, year, council_district, neighborhood) %>%
  summarise(num_crimes = n()) %>%
  mutate(zip_code = as.factor(zip_code))

model1 <- glm(formula = num_crimes ~ neighborhood + council_district + year + zip_code, 
              data = df_agg, family = poisson(link = 'log'))

#summary(model1)
```

Plotting values of coefficients for the various qualitative variables:
``` {r}
length(model1$coefficients)

ggplot() +
  geom_line(aes(x = 1:206, y = model1$coefficients[2:207])) + 
  #scale_x_discrete(labels = names(model1$coefficients[2:207])) +
  scale_x_continuous(breaks = 1:206, labels = names(model1$coefficients[2:207])) +
  xlab("Estimated Neighborhood Coefficients") +
  ylab("Value") + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 2)) 

ggplot() +
  geom_line(aes(x = 1:29, y = model1$coefficients[210:238])) + 
  scale_x_continuous(breaks = 1:29, labels = names(model1$coefficients[210:238])) +
  xlab("Estimated Zip Code Coefficients") +
  ylab("Value") + 
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 8)) 
```


```{r}
model2 <- glm(formula = num_crimes ~ council_district + year + zip_code, 
              data = df_agg, family = poisson(link = 'log'))

#summary(model2)
```
Comparing models, model 1 has a AIC of `r model1$aic` vs model 2 
which has `r model2$aic`. The model with the smaller AIC considered better performing in terms of 
complexity and performance
