---
title: "Simulations"
output: html_document
date: "2024-07-30"
author: "Jason Luo"
---

```{r, message = FALSE, warning=FALSE}
rm(list = ls())

setwd('/Users/jasonluo/Documents/R_crime_analysis')
library(tidyverse)

# Importing cleaned data
df <- read.csv('RMS_Crime_Incidents_Cleaned.csv')
```
### Inter-arrival Times

Inter-arrival times follow an exponential distribution. The lambda parameter is
estimated using the sample mean from our data.
``` {r}

ggplot(data = df, aes(x = interarrival_time)) + 
  geom_histogram(fill = 'darkblue', alpha = 0.6, bins = 30, binwidth = 0.25) + 
  ggtitle("Distribution of Inter-arrival Times of Incidents") + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 

sample_mean <- mean(df$interarrival_time)
lambda <- 1/sample_mean
```
The sample mean is `r sample_mean` and thus the value of lambda is `r lambda`. This
means there are `r sample_mean` hours between incidents and there are on average `r lambda`
incidents per hour.

Since inter-arrival times are exponentially distributed, a sum of inter-arrival 
random variables has a gamma distribution. This sum is the arrival times (till the nth arrival)

``` {r}
n <- 100 # 100 arrivals
set.seed(1234)

generate_arrival_times <- function(n) {
  interarrival_time_samples <- rexp(n = n, rate = lambda)
  arrival_time_sample <- sum(interarrival_time_samples)
  return(arrival_time_sample)
}

arrival_times <- replicate(1000, generate_arrival_times(n)) # generates 1000 samples of arrival times
arrival_times_df <- as.data.frame(arrival_times)
```

``` {r, warning = FALSE}

gamma_samples <- rgamma(n = 1000, shape = n, rate = lambda)
gamma_samples_df <- as.data.frame(gamma_samples)

# Plot 2 ways to simulate fist n arrival times, which follow a gamma distribution
ggplot() + 
  geom_histogram(data = gamma_samples_df, aes(x = gamma_samples), fill = 'purple', alpha = 0.4) + 
  geom_histogram(data = arrival_times_df, aes(x = arrival_times), fill = 'red', alpha = 0.4) +
  ggtitle(paste0("Theoretical Distribution(s) of Arrival Times of First " , n, " Incidents")) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  theme_minimal()

```

We can check how accurate our sample of inter-arrival times follows an exponential distribution
``` {r}
set.seed(12345)
theoretical_sample <- rexp(n = dim(df)[1], rate = lambda)
emperical_sample <- df$interarrival_time

qqplot(emperical_sample, theoretical_sample, main = paste0("QQPlot For Inter-arrival Times of Incidents"))
abline(a = 0, b = 1, col = 'orange')
```

### Simulating a Poisson Process
We estimate the rate parameter for our poisson process by computing
`num_incidents  / total_time `
``` {r}
num_incidents <- nrow(df)
total_time <- as.numeric(difftime(df$incident_timestamp[dim(df)[1]], df$incident_timestamp[1], units = 'hours'))
rate <- num_incidents / total_time

```

There are `r rate` incidents an hour. This is the exact same as the lambda
value from the exponential distribution computed prior. We can now simulate a poisson process:

``` {r}
set.seed(12345)

pois_process <- function(num_arrivals, lambda) {
  arrival_time_stamps <- c() # vector of arrival times
  
  for(i in 1:num_arrivals) {
    inter_arrival_time <- rexp(n = 1, rate = lambda) # inter-arrival times ~ exp(lambda)
    
    if (i == 1) {
      temp <- 0
    } else {
      temp <- arrival_time_stamps[i-1] 
    }
    
    arrival_time <- sum(temp, inter_arrival_time) 
    # sum of N inter-arrival times is the arrival time for the Nth event
    arrival_time_stamps <- c(arrival_time_stamps, arrival_time) 
    # append arrival times to vector
  }
  
  Nt <- 1:length(arrival_time_stamps)
  plot(c(0, Nt) ~ c(0, arrival_time_stamps), type = 's', main = paste0("Simulated " , num_arrivals, " Incidents"),
       xlab = "Arrival Time", ylab = "Number of Incidents")
  
}


for (i in c(5,10,20,50, 100)) {
  pois_process(num_arrivals = i, lambda = rate)
}
```


### Binomial Poisson Model
``` {r}
categories <- unique(df$offense_category)
probs <- c()

for (i in 1:length(categories)) {
  prob_of_category <- sum(df$offense_category == categories[i]) / nrow(df)
  print(paste0("Probability of commiting ", categories[i], " offense: ", prob_of_category))
  probs[i] <- prob_of_category
}
```

Binomial Poisson model for simulating the number of particular incidents

``` {r}
set.seed(123)

binPois <- function(n, prob) {
  y <- rpois(n, lambda = lambda)
  x_y <- rbinom(n, size = y, prob = prob)
  
  x_y_df <- as.data.frame(x_y)
  
  graph <- ggplot(data = x_y_df, aes(x_y)) +
    geom_histogram(bins = 20, alpha = 0.6, fill = "red") +
    theme_minimal() +
    ggtitle(paste0("Estimated Distribution of ", n, " Incidents")) + 
    theme(plot.title = element_text(hjust = 0.5))
  
  return(graph)
}

binPois(n = 1000, prob = probs[1])
```

